---
title: "DR-SC: installation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DR-SC}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Install the DR.SC
This vignette provides an introduction to the R package `DR.SC`, where the function `DR.SC` implements the model `DR-SC`, spatial clustering with hidden Markov random field using empirical Bayes.
The package can be installed with the following command from [Github](https://github.com/feiyoung/ILSE):

`install.packages('remotes')`
`remotes::install_github("feiyoung/DR.SC")`

or install from [CRAN](https://cran.r-project.org/)

`install.packages("DR.SC")`




The package can be loaded with the command:
```{r  eval=TRUE}

library("DR.SC")
```


## Setup on Linux or MacOS system
For running big data, users can use the following system command to set the C_stack unlimited in case of `R Error: C stack usage is too close to the limit`.
```{Linux}
ulimit -s unlimited
```